---
date: 02/09/25
tags:
  - "#Math"
  - "#HW"
links: 
deadline: 2025-02-11
status: 0
format:
  pdf:
    highlight-style: github
---
# 1 
$P(X)=\frac{8}{7}\frac{1}{2}^{X}$

cdf $=F(a)=\sum\limits_{n=1}^{a}f(n)=\sum\limits_{n=1}^{a}\frac{8}{7}\frac{1}{2}^n$

$$
F(x) =
\begin{cases}
0 & \text{for} \quad x < 1 \\
\frac{4}{7} & \text{for} \quad x = 1 \\
\frac{6}{7} & \text{for} \quad x = 2 \\
1 & \text{for} \quad x = 3 \\
\end{cases}
$$

## A $P(X<1.5)$

$P(X<1.5)=F(1)=\frac{4}{7}$

## B $P(X\leq 3)$

$P(X\leq 3)=F(3)=1$

## C $P(X > 2)$

$P(X>2)=F(3)-F(2)=\frac{1}{7}$ 

## D $P(1<X\leq 2)$

$P(1<X\leq 2)=F(2)-F(1)=\frac{2}{7}$

# 2
## A
 
$$
f(X) =
\begin{cases}
0.06 & \text{for} \quad x=350 \\
0.1 & \text{for} \quad x= 450\\
0.47 &\text{for} \quad x= 550 \\
0.37 &\text{for} \quad x= 650
\end{cases}
$$

## B

$\mu = E(X) = \sum\limits_{x}xf(x)=0.06*350+0.1*450+0.47*550+0.37*650=565$

$\sigma = \sqrt{VAR(X)}=\sqrt{E(X^2)-\mu^2}=$
$\sqrt{0.06*350^2+0.1*450^2+0.47*550^2+0.37*650^2-565^2}=\sqrt{6875}=82.9156$

# 3
## A Exactly 5 right

$P(X=5)=C^{25}_{5}(0.25)^5(0.75)^{20}=\frac{25*24*23*22*21}{5*4*3*2}0.25^{5}0.75^{20}=0.16453$

## B More than 2

$P(2<X)=1-P(X\leq 2)=0.96789$

## C Fewer than 3

$P(X<3)=P(X\leq 3)-P(X=3)=0.0321$

## D
```{r}
A <- dbinom(5,25,0.25)
print(paste("A:", A))
print(paste("B:", (1-pbinom(2,25,0.25))))
print(paste("C", (pbinom(3,25,0.25)-dbinom(3,25,0.25))))
```
## E

$\mu=np=6.25$

$\sigma = np(1-p)=4.6875$

# 4
$12$ per hour $=0.2$ per minute, so $\lambda = 0.2,T=$ minutes

## A
5 minutes, so $\lambda T= 1$

$P(X=0)=\frac{e^{-0.2(5)}(0.2*5)^{0}}{1}=\frac{1}{e}=0.36788$

## B

$P(X\geq 3)=1-P(X\leq 3)+P(X=3)=0.0803$

## C
10 mintes, so $\lambda T=2$

$P(X=0)=\frac{e^{-2}(0.2*5)^0}{1}=\frac{1}{e^2}=0.135335$

## D

$P(X=0)=0.001=\frac{e^{-0.2*T}(0.2T)^0}{1}\rightarrow 0.001=e^{-0.2T}\rightarrow 1000=e^{0.2T}=0.2T=\ln(1000)\rightarrow T=5\ln(1000)$

## E
```{r}
print(paste("A:", dpois(0,0.2*5)))
print(paste("B:",1-ppois(3,1)+dpois(3,1)))
```
