---
format: pdf
---

# HW5 STAT 312

## 1
The fill volume of an automated filling machine used for filling cans of carbonated beverage is normally
distributed with a mean of 12.4 fluid ounces and a standard deviation of 0.1 fluid ounce. ($N(12.4, 0.1^2)$)

### A
What is the probability that a fill volume is less than 12 fluid ounces?

> $z=\frac{x-\mu}{\sigma}=\frac{12-12.4}{0.1}=-4$
>
> $P(Z\leq -4)=0.000033$

### B
If all cans less than 12.1 or more than 12.6 ounces are scrapped, what proportion of cans is scrapped?

> $z=\frac{x-\mu}{\sigma}=\frac{12.1-12.4}{0.1}=-3$
>
> $P(Z\leq -3)=0.00135$
>
> $z=\frac{x-\mu}{\sigma}=\frac{12.6-12.4}{0.1}=2$
>
> $P(Z\geq 2)=0.0.02275$
>
> The percent of cans that are scrapped is $0.00135 + 0.02275 = 0.0241=2.41$%
>

### C
Determine specifications that are symmetric about the mean that include 99% of all cans.

> $0.99=P(-z\leq Z\leq z)=P(Z\leq z)-P(Z\leq -z)$
>
> Alternatively, $0.99=\int_{-z}^{z}\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}dx$
>
> This means to find $z$ such that the first statement is true we need to find $z$ such that $P(Z\leq z)=0.995$ and $P(Z\leq -z)=0.005%$. 
>
> Using the z table we find that $z=\pm 2.58$.
>
> We can then use this to find the range by solving $z=\frac{x-\mu}{\sigma}$ for $x$.
>
> $\pm 2.58=\frac{x-12.4}{0.1}\rightarrow x=12.4\pm 2.58\times 0.1=12.658$ or $12.142$
>
> The specifications are $12.142\leq x\leq 12.658$


### D
Use R to find solutions for parts (a) and (c). Attach the R codes and output.

```{r}

print(paste("1A: ", pnorm(12,12.4,0.1)))

print(paste("1C: ", "(",qnorm(0.005,12.4,0.1),",",qnorm(0.995,12.4,0.1), ")"))

```

## 2
The diameter of the dot produced by a printer is normally distributed with a mean diameter of 0.002
inch and a standard deviation of 0.0004 inch.
### (a)
What is the probability that the diameter of a dot exceeds 0.0026?

> $z=\frac{x-\mu}{\sigma}=\frac{0.0026-0.002}{0.0004}=1.5$
> 
> $P(Z> 1.5)=1-P(Z\leq 1.5)=1-0.93319=0.06681$

### (b)
What is the probability that a diameter is between 0.0014 and 0.0026?

> $z=\frac{x-\mu}{\sigma}=\frac{0.0026-0.002}{0.0004}=1.5$
> 
> $z=\frac{x-\mu}{\sigma}=\frac{0.0014-0.002}{0.0004}=-1.5$
> 
> $P(-1.5\leq Z\leq 1.5)=P(Z\leq 1.5)-P(Z\leq -1.5)=0.93319-0.06681=0.86638$

### (c)
What standard deviation of diameters is needed so that the probability in part (b) is 0.995?

> $0.995=P(-z\leq Z\leq z)=P(Z\leq z)-P(Z\leq -z)$ so $0.9975=P(Z\leq z)$ and $0.0025=P(Z\leq -z)$
> 
> Using the Ztable, we find that $z=\pm 2.81$
>
> We can then use this to find the range by solving $z=\frac{x-\mu}{\sigma}$ for $\sigma$.
>
> $\pm 2.81=\frac{\pm 0.0006}{\sigma}\rightarrow \sigma=\frac{0.0006}{2.81}=0.0002135231$


## 3
Suppose that X has a lognormal distribution with parameters $θ = 5$ and $ω^2 = 9$.

### (a)
Use Table III to find P( X < 13,300).

> $P(X<13,300)=P(\ln(X)<\ln(13,300))=P(\ln(X)<9.4955)$
>
> $z=\frac{\ln(13,300)-5}{3}=1.4985$
>
> Using the z-table with $z=1.4985$ we find that $P(\ln(X)<1.4985)=0.933193$.

### (b)
Use Table III to find the value of x such that P( X ≤ x) = 0.95.

> The corresponding z-score for a probability of 0.95 is $z=1.65$.
>
> Thus we need to solve $z=\frac{\ln(x)-5}{3}$ for $x$.
>
> $z=\frac{\ln(x)-5}{3}=1.65\rightarrow \ln(x)=5+3\times 1.65=20952.22$

### (c)
Find the mean and variance of X.

> $\mu=E(X)=e^{\theta+\omega^2/2}=13359.7268$
>
> $\sigma^2=V(X)=e^{2\theta+\omega^2}(e^{\omega^2}-1)=1.446\times 10^{12}$

### (d)
Use R to find solutions for parts (a) and (b). Attach the R codes and output. Please note that your answers in this part may not be exactly the same as your answers in parts a and b when using table III.

```{r}

print(paste("3A: ", plnorm(13300,5,3)))

print(paste("3B: ", qlnorm(0.95,5,3)))

```

## 4
Suppose that X has a lognormal distribution and that the mean and variance of X are 100 and 85,000, respectively.

### (a)
Determine the parameters $θ$ and $ω^2$ of the lognormal distribution. You can use the method that we discussed in class or the hint below to solve the problem.

Hint: define $x= exp( θ )$ and $y= exp( ω^2 )$ and write two equations in terms of x and y.

> $\theta=x\sqrt y=100$, $\omega^2=x^2y(y-1)=85000$
>
> $x\sqrt y=100\rightarrow x^2y=10000\rightarrow x^2y(y-1)=10000(y-1)=85000\rightarrow 10000y=95000\rightarrow y=9.5$
>
> $x\sqrt{9.5}=100\rightarrow x=32.444$
>
> Thus $θ=3.47952$ and $ω^2=2.25129$.

### (b)
Use R to find P$(X\leq 5)$. Attach the R code and output.

```{r}

print(paste("4B: ", plnorm(5,3.47952,2.25129)))

```