---
format: html
---

# Problem 1 (7 pt)
For the hypothesis test $H_0: \mu = 10$ against $H_1: \mu < 10$ with variance unknown and $n = 20$, let the value of the test statistic be $t_0 = 1.25$.

## a. Use table V to approximate the P-value.

> $t_0=1.25$ and we have $20-1=19$ degrees of freedom.
>
> A $t$ value of $1.25$ is between $0.688$ and $1.328$, which results in $0.25<\alpha<0.10$ or $0.75<P<0.9$. Since $1.25$ is a lot closer to $1.328$ than $0.688$ then we can approximate $P\approx 0.88$

## b. Use R to compute the P-value. Attach the code and output.

```{r}
pt(1.25, df=19, lower.tail=TRUE)
```

## c. Does your answer in part b agree with your answer in part a? Why or why not?

> Yes they do because R got an answer of $0.8867613$ which is in between the range specified above and my approximation of $0.88$ is close to the R output.

# P287 8-40 (Modified) (37 pt)
A machine produces metal rods used in an automobile suspension system. A random sample of 15 rods is selected, and the diameter is measured. The resulting data (in millimeters) are as follows: 8.24 8.21 8.23 8.25 8.26 8.23 8.20 8.26 8.19 8.23 8.20 8.28 8.24 8.25 8.24

## a. Check the assumption of normality for rod diameter in the population using a normal probability plot. What is your conclusion and why? Attach the plot.

```{r}
#| ouput: false
#| warning: false
#| include: false

install.packages("qqplotr", repos = "http://cran.us.r-project.org")
library(ggplot2)
library(qqplotr)
```
```{r}
data = c(8.24, 8.21, 8.23, 8.25, 8.26, 8.23, 8.20, 8.26, 8.19, 8.23, 8.20, 8.28, 8.24, 8.25, 8.24)

ggplot(mapping = aes(sample = data)) + stat_qq_point(size = 2,color = "red") + stat_qq_line(color="green") + xlab("x-axis") + ylab("y-axis")
```

> All points are pretty close to the line and there is no clear curve to indicate a skewed distribution. Therefore the population is normal.

## b. Calculate a 95% two-sided confidence interval for the mean rod diameter of all the metal rods produced by this machine. 

```{r}
print(mean(data))
print(sd(data))
```

> $df=15-1=14$, 2 tailed 95% $t$ for $df=14$ is $t=2.145$
>
> $\mu = \bar x \pm t \times \frac{s}{\sqrt n}$
>
> $8.234\pm 2.145\times\frac {0.02529822}{\sqrt {15}}$
>
> $\mu =[8.22, 8.248]$

## c. Based on the data given, can you conclude that the mean rod diameter of all the metal rods produced by this machine is different from 8.25 mm at a significant level of 0.01? Use the 8-step P-value approach.

1. Identify the parameter of interest

> $\mu$ is mean diameter of rods

2. State the hypotheses to be tested.

> $H_0: \mu=8.25$ 
>
> $H_1: \mu \neq 8.25$

3. Specify the level of significance α.

> 2 tailed so $\alpha = 0.005$

4. Determine the test statistic and its distribution.

> $t$ test:
>
> $t=\frac{\bar x-\mu}{s/\sqrt n }= \frac{\bar x - 8.25}{s/\sqrt {15}}$
>
> $t\approx t_{14}$ when $H_0$ is true

5. Calculate the observed value for the test statistic.

> $t=\frac{\bar x-\mu}{s/\sqrt n }= \frac{\bar x - 8.25}{s/\sqrt {15}}$
>
> $= \frac{8.234 -8.25}{0.02529822\sqrt{15}}=-2.4495$

6. Compute the P-value

> $P(t\leq -2.4495)=P(t\geq2.4495)$
>
> $2.145 < 2.4495 < 2.624$
>
> So $0.025 <p< 0.01$

7. Make a statistical decision.

> Fail to reject $H_0$ because $p > 0.005$

8. State the conclusion.

> At the $0.01$ significance level there is not enough evidence that the mean diameter of all rods produced by this machine is different than $8.25$


## d. Use R to conduct the same test given in part c. The resulting output should include all important information that can be used for 3 testing approaches, the rejection region approach, the p-value approach and the confidence interval approach. Attach the code and output.

```{r}
t.test(data, mu=8.25, conf.level =0.99)
```

## e. Can you use the confidence interval in part b to conduct the test in part c? Why or why not?

> No because they have different confidence intervals

## f. Based on the data given, can you conclude that the variance of rod diameter of all the metal rods produced by this machine is greater than 0.0002 mm at a significant level of 0.01? Use the 8-step rejection region approach.

1. Identify the parameter of interest

> $\sigma^2$ is standard deviation of the diameter of all rods in the population

2. State the hypotheses to be tested.

> $H_0: \sigma^2=0.0002$ 
>
> $H_1: \sigma^2 \geq 0.0002$

3. Specify the level of significance α.

> 1 tailed so $\alpha = 0.01$

4. Determine the test statistic and its distribution.

> $\chi^2$ test:
>
> $\chi^2 = (n-1)\frac{s^2}{\sigma^2}= 14\frac{s^2}{0.0002}=70,000 s^2$
>
> $\chi^2\approx \chi^2_{14}$ when $H_0$ is true and when the population distribution is normal

5. Determine the rejection region

> $\chi^2_{0.01,14}=29.14$
>
> If $\chi^2>29.14$ then we reject

6. Compute the $\chi^2$-value

> $\chi^2 = (n-1)\frac{s^2}{\sigma ^2}= 14\frac{s^2}{0.0002^2}=70,000 s^2$
>
> $=70,000 (0.02529822^2)=44.80$

7. Make a statistical decision.

> Reject $H_0$ because $\chi^2 > 29.14$

8.

> At the $0.01$ significance level there is enough evidence that the population variance of the diameter of all rods produced by this machine is greater than 0.0002

## g. Do you need any additional assumptions to justify your test in part f? Why or why not?

> No additional assumptions because normailty was verified in part a

## h. Use an appropriate table to approximate the P- value for the test in part f.

> $44.80 > 31.32$ so $p <0.005$

## i. Use R to calculate the P-value for the test in part f. Attach the code and output.
```{r}
pchisq(44.8, 14, lower.tail=FALSE)
```
