# HW 2 Aidan Bugayong

```{r}
#| ouput: false
#| warning: false

# install.packages('GGally', repos = "http://cran.us.r-project.org")
# install.packages("car", repos = "http://cran.us.r-project.org")
# install.packages("faraway", repos = "http://cran.us.r-project.org")

library(faraway)
library(GGally)
library(car)
```


## 1

```{r}
data = read.csv("data/slumptest-1.csv", sep=",")

dim(data)

head(data)

ggpairs(data)

fit <- lm(CompressiveStrength ~ ., data=data)
summary(fit)
```
### i
According to the above summary, coarse aggregate has a p-value of $0.04546$ which is less than $\alpha =0.05$ so we can say that coarse aggregate is a statistically significant predictor. 

### ii
```{r}
## Compute the residuals
std_resid <- fit$residuals

par(mfrow=c(2,2))
plot(fit, which=2)
hist(rstandard(fit), main="Histogram of standardized residuals")
boxplot(rstandard(fit), main="Boxplot of standardized residuals") 

## Shapiro-Wilk's test
shapiro.test(std_resid)
```


Here the p-value is $0.0467$ which is less than $\alpha =0.05$ so we reject and thus the data is not normally distributed. 

### iii
```{r}
## Cook's distance
D <- cooks.distance(lm(fit))
cutoff1 <- qf(0.5, 8, df.residual(fit))
print(cutoff1)
cutoff2 <- 8/df.residual(fit)
cutoff3 <- 0.5

plot(fit, which=4)
abline(h=cutoff1, lty=2, col="black")
abline(h=cutoff2, lty=2, col="blue")
abline(h=cutoff3, lty=2, col="red3")

influencePlot(fit, id=list(n=3))
```

There are quite a few outliers in the data. These being entries $8,14,49$ and they are not influential because their cook distances are below the cutoff 1 of $~0.925$

### iv 
```{r}
data_no_outliers <- data[-c(8, 14, 49), ]
fit2 <- lm(CompressiveStrength ~ ., data = data_no_outliers)

summary(fit)
summary(fit2)
```

First there is a vast decrease in the intercept from fit1 to fit2. This was compensated by all the other coeffeciants increasing by $~0.01-0.02$. More importantly are the p-values where cement and fly.ash both gained a $*$ (indicating an increase in statistical significance) while coarse aggregate actually decreased in significance. 

## 2

```{r}
data(gala)
dim(gala)

head(gala)

ggpairs(gala)

fit3 = lm(Species ~ ., data = gala)

summary(fit3)
```

### i
```{r}
null_fit = lm(Species ~ 1, data = gala)
anova(fit3, null_fit, test="F")
```

We notice from the anova table that the p-value is much smaller than $\alpha = 0.05$ thus our model is useful.

### ii
From the summary we note that the adjusted $R^2$ is $0.9362$ which means that $93.62%$ of the variance of Species is explained by our predictors. 

### iii
```{r}
vif(fit3)
```

Elevation has a vif above $10$ so there is reason to believe multicollinearity is present.

### iv
```{r}
reduced_model <- lm(Species ~ Area + Elevation, data = gala)
anova(reduced_model, fit3)

```

The claim is fallse because the p-value is less than $\alpha=0.01$ so we reject the null hypothesis thus the predictors are useful.

### v

```{r}
ncvTest(reduced_model)
```

Since the p-value is less than $\alpha=0.05$ the constant variance assumption is violated.
