---
format: pdf
---
```{r}
#| ouput: false
#| warning: false
#| include: false

# install.packages('GGally', repos = "http://cran.us.r-project.org")
# install.packages("car", repos = "http://cran.us.r-project.org")
# install.packages("faraway", repos = "http://cran.us.r-project.org")
# install.packages("onewaytests", repos = "http://cran.us.r-project.org")

library(faraway)
library(GGally)
library(car)
library(onewaytests)
```

```{r}
data = read.csv("data/expenditure.csv", sep=",")

dim(data)

head(data)
```

# i
```{r}
interaction.plot(x.factor = data$selfemp, trace.factor = data$homeowner, response=data$expenditure, xlab="Employment Type (Self Employed)", ylab="Expenditure", trace.label = "Home Ownership", col=c("red","blue"))

```

Since the lines are not parallel, there is most likely an interaction affect.

# ii

```{r}
model_full <- lm(expenditure ~ age + income + homeowner + selfemp + dependents + 
                 homeowner*selfemp, data = data)
summary(model_full)
```


# iii
According to the summary in part ii, the interaction effect coefficient had a p-value of $0.0946$ which is less than $\alpha=0.10$ so the interaction effect is significant.

# iv
I will use a reduced model based on the coefficients that had a p-value less than $0.1$. Thus:

```{r}
model_red <- lm(expenditure ~ age + income + homeowner*selfemp, data = data)
summary(model_red)
```

# v
```{r}
anova(model_full, model_red, test="F")
```

Here the F-test results in a p-val = $0.5335>\alpha=0.1$ so there is no significant relation between expenditure and the extra parameters. 

```{r}
summary(model_red)$r.squared
summary(model_red)$adj.r.squared
plot(model_red, which=1)
```

However, both the $R^2$ and $R_{adj}^2$ are rather small so the model may not be the best fit. Additionally, looking at the residual vs fitted plot has a changing spread which could indicate non constant variance and poor fit.


# vi
```{r}
ncvTest(model_red)
```

Here the p-value is less than any reasonable $\alpha$ so we reject the null hypothesis that there is constant variance f the residuals. We also noted that the constant variance assumption may not be satisfied in part v.

# vii
```{r}
plot(model_red, which=2)
hist(rstandard(model_red), main="Histogram of standardized residuals")
boxplot(rstandard(model_red), main="Boxplot of standardized residuals") 

## Shapiro-Wilk's test
shapiro.test(model_red$residuals)
```

Here the p-value is less than any reasonable $\alpha$ so we reject the null hypothesis that the residuals are normally distributed.

It is also clear from the QQ plot and the box plot that this distribution is likely not normal.

# viii
```{r}
durbinWatsonTest(model_red)
```

Here the p-value $(0.946)$ is greater than any reasonable $\alpha$ so we do not reject the null hypothesis and the independence assumption is satisfied.

# ix
```{r}
## Cook's distance
D <- cooks.distance(lm(model_red))
cutoff1 <- qf(0.5, 4, df.residual(model_red))
print(cutoff1)
cutoff2 <- 4/df.residual(model_red)
cutoff3 <- 0.5

plot(model_red, which=4)
abline(h=cutoff1, lty=2, col="black")
abline(h=cutoff2, lty=2, col="blue")
abline(h=cutoff3, lty=2, col="red3")
## influence plot
influencePlot(model_red)
```

$F(0.5,4,998)=0.8397453$ and none of the points have a cooks distance above this point. Therefore none of the points are influential (although we have a few outliers.)

# x
```{r}
vif(model_red)
```

Here all the VIF scores are less than 10 so we do not believe there is any multicolinearity.
